Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-10 second-draft/pnas-word-order-si.tex
SI

\usepackage{natbib}

@@

@@

\title{Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''}
\author{
        Michael Hahn \\
                Department of Linguistics\\
       Stanford University
            \and
       Daniel Jurafsky\\
       Department of Linguistics\\
       Stanford University \\
       \and
       Richard Futrell\\
       Department of Language Science\\
       University of California, Irvine
}
\date{\today}

@@

\textcolor{red}{This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}

\section{Study 2: Results for Individual Languages}

\section{Study 2: Regression for Predicted Correlations}

@@

@@

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding recent objections by~\cite{dunn2011evolved}, see follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's seventeen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

\section{Study 1: Supporting Analyses for Study 1}

@@

\section{Study 2: Results on all UD Relations}
In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
We also report results for grammars optimized for predictability and parseability individually.

@@

\section{Study 2: Comparing Efficiency to its Components}\label{sec:posterior-number}

@@

%\section{Joint optimization of grammar, parser, and language model parameters}
\section{Optimizing Grammars for Efficiency}

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
Our main methodological innovation is to carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

@@

\textcolor{red}{do same for language model, and for the other random grammars}

The resulting curves are shown in Figure~\ref{fig:learning-czech}.
A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.

\begin{figure}
    \centering
    \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.\textcolor{red}{font size}}
    \label{fig:learning-czech}
\end{figure}

\section{Study 2, Correlation between Universals and Efficiency}

>>>>

\usepackage[numbers]{natbib}

@@

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

@@

\title{Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''}
\author{
        Michael Hahn \\
                Department of Linguistics\\
       Stanford University
            \and
       Daniel Jurafsky\\
       Department of Linguistics\\
       Stanford University \\
       \and
       Richard Futrell\\
       Department of Language Science\\
       University of California, Irvine
}
%\author{
%Michael Hahn, Daniel Jurafsky, Richard Futrell}

%\correspondingauthor{Michael Hahn.\\E-mail: mhahn2@stanford.edu}

@@

\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}

@@

\textcolor{red}{TODO At some point say something about \cite{dunn2011evolved}, since that got a lot of attention? can refer to follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}

@@

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's seventeen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

\section{Supplementary Analyses for Study 1}

@@

\section{Supplementary Analyses for Study 2}
\subsection{Correlation between Universals and Efficiency}

\subsection{Predictions for Individual Languages}

\subsection{Regression for Predicted Correlations}

\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}

\subsection{Results on all UD Relations}
In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
We also report results for grammars optimized for predictability and parseability individually.

@@

@@

%\section{Joint optimization of grammar, parser, and language model parameters}
\section{Creating Optimized Grammars}

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

@@

The resulting curves are shown in Figure~\ref{fig:learning-czech}.
%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.

\begin{figure}[ht]
    \centering
%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}

        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
    \label{fig:learning-czech}
\end{figure}
